---
title: "Lecture_9_Walkthrough"
author: "LQ"
date: "5/1/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
#View(wisc.df)
head(wisc.df)
```

```{r}
#Table uses the cross-classifying factors to build a contingency table of the counts at each combination of factor levels; $calls the diagnosis column from the dataframe
nrow(wisc.df)
table(wisc.df$diagnosis)

```

```{r}
colnames(wisc.df)

length(grep(pattern = "_mean", colnames(wisc.df)))


```

###Q5. Why do you think we are using the indices 3:32 here?
####Because the id and diagnosis columns will not be used & the last column is NA (when the data went through excel, a space was accidentally input at some point yielding the weird X column filled with NA)

```{r}
#Select cols 3:32 and store as a matrix
wisc.data <- as.matrix(wisc.df[,3:32])
#View(wisc.data)
```


```{r}
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
head(wisc.data)

diagnosis <- wisc.df$diagnosis
```


##PCA
```{r}
colMeans(wisc.data)

#apply(data,columns - 1 is rows, function)
apply(wisc.data,2,sd)
round(apply(wisc.data,2,mean), 1)

```


```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
# Look at summary of results
#summary(wisc.pr)
plotwisc.pr <- biplot(wisc.pr)
```


###Q10. What stands out to you about this plot? Is it easy or difficult to understand? Why?
####Eveything is too clumped
```{r}
# Scatter plot observations by components 1 and 2
diagnosis <- wisc.df$diagnosis

plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, 
     xlab = "PC1", ylab = "PC2")

```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)

# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)
pve

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Percent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


####install.packages("factoextra")
```{r}
library("ggplot2")
library("factoextra")
## ggplot based graph
fviz_eig(wisc.pr, addlabels = TRUE)

```


```{r}
wisc.pr$rotation[,1]
```


```{r}
wisc.pr$rotation["concave.points_mean",1]
sort( abs(wisc.pr$rotation[,1]) )
```

###Hierarchical clustering
####hclust(dist(x))
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data, scale = TRUE)
head(data.scaled)

```

```{r}
data.dist <- dist(data.scaled, method = "euclidean")
wisc.hclust <- hclust(data.dist, method = "complete" )
wisc.hclust
plot(wisc.hclust)
abline(h = 19, col = "red", lty = 2)
```


###One-liner:
```{r}
wisc.hclust <- hclust(dist(scale(wisc.data)))
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
##Can also cut at h = 19, k is the number of clusters
table(wisc.hclust.clusters)
table(wisc.hclust.clusters, diagnosis)
```
####We don't need a readout for the individual arguments so it is called within one function

###Combining PCA with hclust
```{r}

#90% capture or original variance in the dataset - PC1 - PC7
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2")
plot(wisc.pr.hclust)
abline(h = 70, col = "red", lty = 2)

```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = grps)
```
###install.packages("rgl")

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)
```

##SECTION 7 - PREDICTIONS
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
g <- as.factor(grps)
levels(g)
g <- relevel(g,2)
levels(g)
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```

```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```




